---
title: | 
  | SME. 4th Laboratory 
  |
  | Bivariate Data. Linear Regression.
  |
  |
author: '**Name:** Silvia Arenales Muñoz '
date: |
  |
  | `r Sys.setlocale("LC_TIME", "C");format(Sys.Date(), "%B %d, %Y")`
  |
  | Table of contents
  |
output:
  html_document:
    highlight: espresso
    number_sections: yes
    theme: united
    toc: yes
    toc_depth: 4
  pdf_document:
    toc: yes
    toc_depth: '4'
    keep_tex: true
---




```{r setup, include=FALSE}
# setting the options

knitr::opts_chunk$set(echo = TRUE,tidy=TRUE)
options() # printing all current options
options(digits = 6) # number of digits to display

library(moments)   # needed for finding the third and fourth  moments
# Install if not already installed

```

>     
# Describing relationships  with graphs and numbers

>    

```{r eval=FALSE}
# Load the dataset if you are working on this file alone.

cars <- read.csv("cars_mileage.csv",header=TRUE,sep = ",", stringsAsFactor=FALSE)

TwoS <- cars[cars$Type=="T",]

```



>   

## Graphical display of the relationship. Scatterplot

>   

> Draw a **scatterplot** of highway mileage $Y$ against
city mileage $X$  using only  the two-seater cars.   

>  

```{r }
### Useful commands and functions: plot and all that you know 
### Write your commands below:

names(TwoS)

plot(TwoS$City,TwoS$Hwy,
     main =  "Scatter plot",
     xlab = "City",
     ylab = "Hwy")


```


>  

> Describe  the graphic: 

>* Can you see any relationship? Describe it.  

> **Answer:** Upper rigth, in the corner.  


>* In the previous laboratory we saw that the Insight was an outlier,
can you picture it in the graphic?  

> **Answer:** It falls.


>* What happens when you drop the outlier?   

> **Answer:** ...It falls.

```{r fig.height=10}
# Draw both graphics side by side
# On the left the graphic with the outlier and on the right the graphic without the outlier

### Useful commands and functions: plot, cor and all that you know 
### Write your commands below:


ind <- which(TwoS$City==max(TwoS$City), 
      arr.ind = TRUE)

layout(c(1,2),1,2) #one beside the other

plot(TwoS$City[-ind],TwoS$Hwy[-ind],
     main = "Scanner plot",
     xlab ="city",
     ylab ="Hwy")

plot(TwoS$City,TwoS$Hwy,
     main ="Scatter plot without the outlier",
     xlab = "City",
     ylab = "Hwy")

```


>  

## Measuring the strength of the relationship. Pearson's correlation coefficient

>  

```{r }
### Useful commands and functions:  cov, cor, pairs and all that you know 
### Write your commands below:

cor(TwoS[,2:3])
cor(TwoS[-ind,2:3])#without

pairs(TwoS[-ind,2:3])

```



> Find the correlation between city and highway mileages both with and
without the outlier. Is there any change?   Explain what happens and why.

> **Answer:** Yes,


> Which value is greater? Explain your answer.    

> **Answer:** The one that it is with it.

>  

## Model of the relationship. Linear regression {.tabset .tabset-pills}


>  Find the **regression line** for predicting highway mileage
from city mileage. Find one model with the outlier and other without it.   


### Model without the outlier 

> Find the regression line  for predicting highway milage from city mileage.   

```{r eval=FALSE}
### Useful commands and functions: lm, summary  and all that you know 
### Write your commands below:

RL.without <- RL <- lm(TwoS$Hwy[-ind] ~ 
                         TwoS$City[-ind])   ##Y=aX+b

```


> **Examining the Outputs of a Linear Regression**     

> Before accepting the result of a linear regression it is important to evaluate its 
suitability at explaining the data.   

> One way is to examine the **Variability explained by the model**

> The square of the correlation, $r^{2}$, is the fraction of the variation in the values of $Y$ that is explained by the least-squares regression of $Y$ on $X$.

> $r^{2} = \frac{\displaystyle s^{2}_{\hat{Y}}}{\displaystyle s^{2}_{Y}} = \Sexpr{r.2}$   

> 

> Command *summary*   


```{r eval=FALSE}
### Useful commands and functions: lm, summary  and all that you know 
### Write your commands below:

# multiple-R-squared value?

summary(RL)

coef(RL)

resid(RL)

predict(RL)




```

> 



> Command *plot*   

```{r eval=FALSE}
### Useful commands and functions: lm, summary  and all that you know 
### Write your commands below:

par(mfrow=c(1,1))

plot(RL)

```



### Model with the outlier 

> Find the regression line   for predicting highway milage from city mileage.   


```{r }
### Useful commands and functions: lm and all that you know 
### Write your commands below:

RL <- lm(TwoS$Hwy[-ind] ~  TwoS$City[-ind])   ##Y=aX+b

```



>  Draw both lines in the same graphic. Is there any difference? 

```{r }
### Useful commands and functions: plot, cor and all that you know 
### Write your commands below:

plot(RL)

```




### Comparing both models 

> Find the predicted value for observation number 4 using both models.    


```{r }
### Useful commands and functions: plot, cor and all that you know 
### Write your commands below:

predict(RL)

```


> Find the residual value for observation number 4 using both models.     



```{r }
### Useful commands and functions: plot, cor and all that you know 
### Write your commands below:

resid(RL)
```




