---
title: |
  | LAB 1  (**SME**)
  | Introduction to RStudio and  Rmarkdown 
  |
author: '**Name:** Silvia Arenales'
date: |
  | `r Sys.setlocale("LC_TIME", "C");format(Sys.Date(), "%B %d, %Y")`
  |
  |
output:
  html_document:
    number_sections: yes
    theme: united
    highlight: espresso
    toc: yes
    toc_depth: 3
  pdf_document:
    toc: yes
    toc_depth: '3'
---

>  

>   

```{r setup, include=FALSE}
# R commands inside chunck

# Name of the chunck: setup
# 

# setting global options,
# all the following chunks will be affected by these options
knitr::opts_chunk$set(echo = TRUE,tidy=TRUE)
# echo = TRUE, By default shows all the commands written inside the chuncks
# tidy=TRUE, Whether to reformat code in a tidy way when displaying it


```

>    

>  
# Installing R and RStudio(in that order)  

> Even if you use RStudio, you’ll still need to download R to your computer. RStudio helps you use the version of R that lives on your computer, but it doesn’t come with a version of R on its own.   

> You can read how to do it in the following link: 
[Hands-On Programming with R](https://rstudio-education.github.io/hopr/starting.html)

>    

# Rmarkdown and RStudio

> This file you are working on is written in Rmarkdown(.Rmd) format because I want to write not only R commands but add some extra explanations to them. 

> R Markdown is a format for writing reproducible, dynamic reports with R. Use it to
embed R code and results into slideshows, pdfs, html documents, Word files and more. 

> You can find how to work with R markdown in RStudio in the following file:
[R Markdown Cheat Sheet](rmarkdown-cheatsheet.pdf){width=65%}


> Look in that file for the answers of some questions about the .Rmd file you are working on:   

> **QUESTION: When you compile this file, which is the format of the output file by default?**   
>**ANSWER: ** #The format of the output file by default is R Markdown syntax.

> **QUESTION: What do you need to do in order to include and execute R code in this file?** 
>**ANSWER: ** #Embed R code that created output to include in the report and then replace ir with its output and transform the report into slideshow, pdf, html...

> **QUESTION: If you only want to show the results not the code which parameter do you use in the corresponding chunk of code?** 
>**ANSWER:** #You need to trandoem the report into a slideshow, pdf, html...

> **QUESTION: Which is the extension of this file?** 
>**ANSWER:**  #Pdf 


# Basic commands in R or RStudio

> You can review the slides of the chapter to see how to do it in RStudio using the menus in the upper pannel.

> **Session>Set working directory>Choose Directory**  

>   

## Setting the working directory

> If you want to know in which directory your files are saved:  

```{r echo=TRUE, eval=TRUE, message=FALSE}
# R commands inside chunck

### Useful commands: getwd,setwd

# Assign to the variable myWD your working directory
### Write your commands where needed
#myWD <- setwd("C:/Users/XXXX/Dropbox/informatica_segundo/sme/lab/lab1") # changed DONE 
myWD <- NA
# See what happens when you change that value
## NOTICE that getwd() returns the current wd as a string, but setwd() sets wd and returns the value as a string.

```

>  

> **QUESTION: In the header of the previous chunk there are some paremeters, can you explain their use?**   
>**ANSWER:** #echo means whether to display code along with its results; eval whether to evaluate the code and include its results and message it is used to display messages.



> **QUESTION: What is my working directory? **   
> **ANSWER:** My working directory is `r myWD`

>    

## Information about the system and version of the program.

>  

```{r echo = FALSE}
# R commands inside chunck

### Useful command:

sessionInfo()

```

> 

## Installing and charging the libraries in memory. 

> If a library is not already installed we must install it first. We will install the package corresponding to the 
book "Using R for Introductory Statistics".  

```{r, echo=TRUE, eval=TRUE, tidy=TRUE, message=FALSE}
# R commands inside chunck

### Commands to use: install, help
### See the menu on the Packages window, the result of the actions in the menu will appear in the Console panel.

### install.packages("UsingR")

### NOTICE: the line above is commented because "UsingR" package was already installed succesfully.

```

>  **QUESTION: Who is the author of the book?**    
>  **ANSWER:** The author of the book is `r packageDescription("UsingR")[4]`.

> If the dataset is already included in a package, you only need to load that package.   
  Load the package for the book "Using R for introductory Statistics" and have a look at the datasets included.   
  Choose the dataset named "math"  and read its description in the help menu. 


```{r echo=TRUE, eval=TRUE, message=FALSE, tidy=TRUE}
# R commands inside chunck

### NOTICE: CHANGE THE VALUE OF PARAMETER eval BEFORE COMPILING

### Commands to use: library, data, help
### Write your commands where needed:

library("UsingR")             # load package for the book "Using R for introductory Statistics"
data(package="UsingR")        # show only the datasets included in that library
data(math,package="UsingR")   # load the dataset math in memory

# which command? # look at the explanation about the math dataset in the help window
  # help(data(math,package="UsingR"))

### NOTICE that we commented the line above because otherwise a new .html would be opened on the default internet browser.

# which command? # show the values of the dataset
  math
  
```

>  


# Exercise {#statm}

>  

>  In case you have any doubt, you can look in the help menu or read the first two subsections of the section "The very basics" in the previous link: [Hands-On Programming with R](https://rstudio-education.github.io/hopr/starting.html)



>Your cell-phone bill varies from month to month. The monthly amounts in
dollars for the last year were    
46 33 39 37 46 30 48 32 49 35 30 48



>* 1. Enter this data into a variable called bill.  

bill <- scan(text = "46 33 39 37 46 30 48 32 49 35 30 48")

##bill<- {46; 33; 39; 37; 46; 30; 48; 32; 49 ; 35; 30; 48}

>* 2. Use the sum function to find the amount you
spent last year on the cell phone. 

total<- sum(bill)

>* 3. Associate data with their corresponding month  

January <- bill[1]
February<- bill[2]
March <- bill[3]
April<-bill[4]
May<- bill[5]
June<-bill[6]
July<-bill[7]
August<-bill[8]
September<-bill[9]
October<-bill[10]
November<-bill[11]
December<-bill[12]


>* 4. What is the smallest amount you spent in a month? 

# 33

>* 5. That minimum corresponds to which month? 

# November

>* 6. What is the greatest amount you spent in a month? 

# 49

>* 7. That maximum corresponds to which month? 

# September

>* 8. How many data you have?

total


```{r echo=TRUE,eval=TRUE, tidy=TRUE, message=FALSE,results='asis',hold}
# R commands inside chunck

# (1) scan 
bill <- scan(text = "46 33 39 37 46 30 48 32 49 35 30 48") # inline use
cat("1. Last year's **bills** by month",bill,"\n")
# 
# (2) sum
totalA <- sum(bill) # year's total amount
cat("2. **Total amount** spent in the phone", totalA,"\n")

# Value of the array  in sixth position 
bill.6 <- bill[6]

```

>      

> We can also write the answers outside the code chunk using the results we already have:  
 The value of the bill in sixth position  is `r bill.6`      


>  

> **We can also create a table with our data**      

>   
```{r echo=TRUE,eval=TRUE, tidy=TRUE, message=FALSE,results='asis'}
# R commands inside chunck
#
#(3) names,  naming data(column)
library("kableExtra")
names(bill) <- c("jan","feb","mar","apr","may","june","july","aug","sept","oct","nov","dic")
print("Last year's **bills by month:** ")

kable(t(bill))%>%
  kable_styling()


```


>    


```{r echo=TRUE,eval=TRUE, tidy=TRUE, message=FALSE,results='asis'}
# R commands inside chunck
#
# (4) min
Min.bill <- min(bill)
cat("4. **Minimum** amount of a bill is ",Min.bill,"euros \n")
#
# (5) which
Position.of.min <- which(bill==min(bill),arr.ind = TRUE)
cat("5. Month(s) of minimum spent: ",names(Position.of.min))
```

>    


```{r echo=TRUE,eval=TRUE, tidy=TRUE, message=FALSE,results='asis'}
# R commands inside chunck
#
# (6) max
Max.bill <- max(bill)
cat("6. **Maximum** amount of a bill is ",Max.bill,"euros \n")
#
# (7) which
Position.of.max <- which(bill==max(bill),arr.ind = TRUE)
cat("7. Month(s) of maximum spent: ",names(Position.of.max))
```


>   

```{r}
# R commands inside chunck

# (8) lenght


```


>    

## Defining functions

> As in any other programming languages you can also define functions in R.  

> You will find some useful examples in the sections 2.3,2.4 and 2.5  in the previous link: [Hands-On Programming with R](https://rstudio-education.github.io/hopr/starting.html)

> You can find it resumed in the following picture:   
[Hands-On Programming with R. Parts of an R function](hopr_0106.png){width=25%}

>   


```{r echo=TRUE,eval=TRUE, tidy=TRUE, message=FALSE,results='asis'}
# R commands inside chunck

# Define a function to find the position of maximum value in an array




```


>    


# Saving your work space   

> Click on the  icon of the floppy disk   
Choose a name for the file and save.   


>  **QUESTION: Which is the extension of the file?**    
> **ANSWER:** #It is the extension .Rmd




# Recovering the last commands: Rhistory

> If you want to recover the last commands or save them you can do it in the upper right window using the History option.  

> Remember to clear all  entries in the History.   



